---
title: "Independent Project: "
author: "Minwoo Kang"
date: "December 17, 2022"
output:
  pdf_document: default
  html_document: default
---


```{r knit-settings, include=FALSE, echo=FALSE, warning=FALSE}
# Prevent code from showing up in final report
knitr::opts_chunk$set(echo = FALSE)
```

```{r random-seed, echo=FALSE, warning=FALSE}
set.seed(440)
```

```{r libraries, include=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(eha)
library(survival)
library(survminer)
library(knitr)
library(broom)
library(kableExtra)
library(dplyr)
library(sjmisc)
library(stringr)
```

# 1. Introduction

## 1.1 Background Information on Current Hip-Hop Music Industry

## 1.2 Dataset Used

## 1.3 Variables of Interest

```{r preprocessing}
# Only include patients enrolled in the follow up protocol
data <- rawdata %>% 
  filter(FUPROTCL == 1)
  
# Keep only the first catheterization per patient
data <- data %>% 
  group_by(RSUBJID) %>%
  mutate(NUMCATHS = max(RSEQCATHNUM)) %>%
  filter(RSEQCATHNUM == 1) %>% 
  ungroup()
```

```{r}
billboard <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/billboard.csv')
audio_features <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/audio_features.csv')
```
```{r preprocessing}
# Only include patients enrolled in the follow up protocol
data <- rawdata %>% 
  filter(FUPROTCL == 1)
  
# Keep only the first catheterization per patient
data <- data %>% 
  group_by(RSUBJID) %>%
  mutate(NUMCATHS = max(RSEQCATHNUM)) %>%
  filter(RSEQCATHNUM == 1) %>% 
  ungroup()
```

```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```

```{r}
billboard_cleaned <- billboard %>%
  group_by(song_id) %>%
  arrange(weeks_on_chart) %>%
  slice(n()) %>%
  ungroup() %>%
  select(song_id, peak_position, weeks_on_chart)

audio_features_cleaned <- audio_features %>%
  select(-spotify_track_preview_url, -spotify_track_album)

billboard_cleaned
```

```{r}
audio_features_cleaned <- audio_features %>%
  select(-spotify_track_preview_url, -spotify_track_album)

audio_features_cleaned
```

```{r}
all_data <- merge(x=billboard_cleaned, y=audio_features_cleaned, by="song_id")

all_data
```

```{r}
all_data_hiphop <- all_data %>%
  filter(grepl("hip hop", spotify_genre) == TRUE)

all_data_hiphop
```
Let's use this for now
```{r}
spotify_songs_hiphop <- spotify_songs %>%
  filter(playlist_genre == "rap")

spotify_songs_hiphop
```
```{r}
spotify_songs
```

```{r}
spotify_songs <- drop_na(spotify_songs) %>%
  filter(track_popularity > 0)
spotify_songs
```

# 2. Exploratory Data Analysis (EDA)

track_popularity ~ playlist_genre, danceability, energy, key, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms

```{r}
# Figure for popularity

ggplot(spotify_songs, aes(x = track_popularity)) +
  geom_histogram(colour = 4, fill = "white", bins = 10) +
  labs(x = "Track Popularity",
       y = "Number of Tracks") +
  theme_minimal() +
  theme(axis.title = element_text(size = 8),
        title = element_text(size = 8),
        axis.text = element_text(size = 8))
```

# 3. Methodology

## 3.1 Description of Methodology

## 3.2 Model Specification

## 3.3 Model Assumptions

Missing data?

# 4. Result

# 5. Discussion

## 5.1 Conclusion

## 5.2 Potential Limitations and Future Directions

## 5.3 Summary

# Appendix A

